Dual host-parasite transcriptomes of apicomplexan Eimeria falciformis and its natural mouse host

### 2016-07-21 figuresANDmanuscript/manuscriptBMCformat/manuscript_rnaseq.tex
some work on manuscript

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental overview}
We performed mRNA sequencing of caecum tissue of mice infected with
the parasite \textit{E. falciformis} at different timepoints and for
infections of different mouse strains.

Replicate average of transcripts as order of magnitude for
\textit{E. falciformis}/mouse are indicated in Table.... Columns in
table represent mouse strains used in infection experiments. Rows
represent timepoints post infection plus oocyst and sporozoite
samples. The upper part of the table shows data for first infection,
and oocyst and sporozoite data. The lower part shows data for
challenge infection. Averages were calculated after sample exclusions
(see Methods). For exact values, see Table 1.

@@

Table 2 lists pairwise sample comparisons and number of genes with
differently abundant mRNAs per comparison."NMRI" followed by number
indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3
p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as
implemented in edgeR are included. NAs are missing samples or not
applicable for the species. NA* is due to missing NMRI day 0 sample
from first infection. For \textit{E. falciformis} no mRNAs are
significantly differently abundant between first and second infection,
whereas in mouse there are some differences, especially on day 7
p.i. However, upon hierarchical clustering of these genes (union of
500 with lowest FDR), samples do not cluster according to first/second
infection (Figure SX ?). In comparisons between mouse strains, 22
genes are detected to be different in the parasite data. The
difference is however not between Rag1-/- mice as expected, but
between NMRI and C57BL/6. For mouse....

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We applied hierarchical clustering on \textit{E. falciformis} mRNAs which were differently abundant at different timepoints p.i., including comparisons with oocysts and sporozoites (union of 500 with lowest FDR). Samples from day 7 p.i. form a separate cluster (NMRI mice only), not distinguishing between first and challenge infection. For these samples, the pattern is most pronounced in gene cluster 1 (up), and 4 and 6 (down). Within cluster 1, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 2). Distinct groups of genes also define sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days 3 and 5 p.i. from all three mouse strains cluster together. These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites. Sporozoites cluster closer with days 3 and 5 than with oocysts or day 7.

@@

\section*{Additional Files}
  \subsection*{Additional file 1 --- Raw and normalized counts }
  Raw and normalized counts for individual samples in our
  study (two compressed csv files).

  \subsection*{Additional file 2 --- Results of statistical tests (edgeR) }
  Fold-changes, p-values and false discovery rates (adjusted p-values)
  are given for all tested contrasts (one compressed csv file).

  \subsection*{Additional file 3 --- Additional methods and resuts }
  % this would be cool but requires special referencing in the text
  % (hard to keep track and hard to fit in the journals referencing
  % system). But we can try.
  Document containing additonal figures and summary tables (pdf).

\subsection*{Additional file 4 --- Results of enrichment analyses (topGO)}
  Tables listing all tested gene sets and resulting significatn GO
  terms.

>>>>

We found xx E. falciformis genes of the total predicted yy, which were expressed in at leat one developmental stage measured.
Of the total reported mouse genes aaa, hhh were expressed on at least one timepoint during E. falciformis infection and fff were expressed in controls.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% 	OVERVIEW READS PER SAMPLE, LAYOT AS EXPERIMENTAL OVERVIEW
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental overview}
Replicate average of transcripts as order of magnitude for \textit{E. falciformis}/mouse are indicated in Table.... Columns in table represent mouse strains used in infection experiments. Rows represent timepoints post infection plus oocyst and sporozoite samples. The upper part of the table shows data for first infection, and oocyst and sporozoite data. The lower part shows data for challenge infection. Averages were calculated after sample exclusions (see Methods). For exact values, see Table 1.

@@

Table 2 lists pairwise sample comparisons and number of genes with differently abundant mRNAs per comparison."NMRI" followed by number indicates day post infection (e.g. NMRI 3 = NMRI mouse on day 3 p.i.). Genes with Benjamini-Hochberg corrected p-values $\leq$0.01 as implemented in edgeR are included. NAs are missing samples or not applicable for the species. NA* is due to missing NMRI day 0 sample from first infection. For \textit{E. falciformis} no mRNAs are significantly differently abundant between first and second infection, whereas in mouse there are some differences, especially on day 7 p.i. However, upon hierarchical clustering of these mouse genes (union of 500 with lowest FDR), samples do not cluster according to first/second infection (Figure SX ?). In comparisons between mouse strains, 22 \textit{E. falciformis} genes are detected as different. The difference is however not between Rag1-/- immune incompetent and either WT mouse as expected, but between NMRI and C57BL/6. For mouse.... 
The comparisons between timepoints p.i. are dominant in number of mRNAs being differently abundant and we were interested in finding patterns in this data. Unsupervised hierarchical clustering was applied on \textit{E. falciformis} mRNAs which were differently abundant at different timepoints p.i., including comparisons with oocysts and sporozoites (union of 500 with lowest FDR, Figure 1). Samples from day 7 p.i. form a separate cluster (NMRI mice only), not distinguishing between first and challenge infection. For these samples, the pattern is most pronounced in gene cluster 1 (up), and 4 and 6 (down). Within cluster 1, the second replicate from the challenge infection has a deviant profile (compare mouse data in Figure 2). Distinct groups of genes also define sporozoites (cluster 5 up) and oocysts (cluster 3 up, cluster 6 down). mRNA profiles on days 3 and 5 p.i. from all three mouse strains cluster together. These samples are distinct from oocysts, NMRI day 7 p.i., and sporozoites. Sporozoites cluster closer with days 3 and 5 than with oocysts or day 7.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.
